### 2013-10-25 src2/thesis.tex
update

In 2006 NVIDIA released their first Compute Unified Device Architechture (CUDA) 
GPU. With CUDA, NVIDIA also made it a lot easier for programmers to use GPUs 
for general purpose workloads. A dialect of C, called CUDA C, was released specifically 
to simplify implementation of general purpose algorithms on GPUs.

@@

Papers B, C, E and F (sections~\ref{sec:paperB},~\ref{sec:paperC},~\ref{sec:paperE},
~\ref{sec:paperF}) contain more information about GPUs, and the assumptions we make 
about them in relation to Obsdian.

@@

This kernel takes two input arrays, {\tt i1} and {\tt i2}, and computes 
elementwise sums into the result array, {\tt r}. The {\tt threadIdx} variable identifies 
a thread within a block. There is also a {\tt blockIdx} variable that identifies 
the block a thread and {\tt blockDim.x} specifies the number of threads per block.

@@

>>>>

In 2006 NVIDIA released their first Compute Unified Device Architecture (CUDA) 
GPU. With CUDA, NVIDIA also made it a lot easier for programmers to use GPUs 
for general purpose workloads. A dialect of C, called CUDA C, was released specifically 
to simplify implementation of general purpose algorithms on GPUs.

@@

Papers B, C, E and F (sections~\ref{sec:paperB},~\ref{sec:paperC},~\ref{sec:paperE},
~\ref{sec:paperF}) contain more information about GPUs, and the assumptions we make 
about them in relation to Obsidian.

@@

This kernel takes two input arrays, {\tt i1} and {\tt i2}, and computes 
element wise sums into the result array, {\tt r}. The {\tt threadIdx} variable identifies 
a thread within a block. There is also a {\tt blockIdx} variable that identifies 
the block a thread and {\tt blockDim.x} specifies the number of threads per block.

@@

\subsection{Writing efficient GPU kernels}

Writing efficient GPU kernels is hard. There are many details that the CUDA 
programmer must be aware of in order to make optimal use of the GPU. 
the NVIDIA ``CUDA C Best Practices Guide''~\citet{BestPrac}, is a good source 
to learn more about these details. Here is a list with explanations of some 
of the practices that the guide deem most important.

\begin{itemize}

\item {\bf Minimise data transfers between host and device:} The bandwidth between device memory 
and GPU is over 20 times of the bandwidth between host memory and the GPU device via the PCIe bus.
This means that, first of all one must evaluate if the computation to be perform is significant 
enough to warrant a transfer to the GPU. If it computation is not arithmetic intensive it may be 
better to do the work on the CPU. However, if the data is already in device memory and 
the operation you are about to perform on it is more efficiently handled by the CPU, it may still 
be better to let the GPU do the work. In other words, compute near the data unless transferring it 
pays off.

\item {\bf Ensure that global memory accesses are coalesced:} Certain access patterns into 
GPU device memory can be {\em coalesced} (combined into few memory transactions). Unfortunately 
what these access patterns are very some depending on the GPU. For a representative model of 
GPU the rule is that concurrent memory accesses (by threads within a warp) will be coalesced 
to as many transactions as cache lines touched (128b l1 cache lines). This means that strided 
accesses from within a warp are not favoured.

\item {\bf Minimise use of global memory:} Accessing global device memory can take as 
many 400 to 600 clock cycles if the data is not cached. The programmer should 
use shared memory as much as possible to avoid redundant loading from global memory.
Shared memory could also be used to coalesce reads from global memory. Shared memory is 
a limited resource in the MP, so if a kernel uses very much shared memory it means 
fewer such blocks can be active on the MP.

\item {\bf Use a multiple of 32 threads per block:} Groups of 32 threads, warps, are the 
scheduled unit on a GPU. If a warp accesses memory (and will be waiting for it arrive during 
600 cycles) it will be swapped out and another warp will take its place. This means that 
the number of available warps decides how well latency can be hidden. A block that is not a 
multiple of 32 leads to wasting GPU resources, some processing elements will stand idle during 
the execution of the not full warps.

\item {\bf Avoid different execution paths within a block:} Any branching instructions that 
causes the execution path diverge within a warp will affect performance. When the execution 
paths within a warp diverges the computation is serialised.

\item {\bf Do not use {\tt \_\_synchthreads()} in diverging code:} Extreme care must be taken 
when synchronising within any conditionally executed code; every thread must reach this barrier.
Failing to ensure this will likely lead to the kernel producing incorrect answers. Note that 
this is a major limitation to the composability of CUDA functions. If it is safe or 
not to call some function at a certain place in your code, can only be decided by inspecting 
that functions implementation. 
\end{itemize}

The list above provides some very general guide lines of what to think about when 
writing GPU kernels. The lesson is that fast memory near the processing element 
should be preferred over slow memory far away. However, this comes at the effort 
of decomposing computation and data in a way that makes this possible.

%\citet{FERMIOPT}
%\citet{BestPrac} 
%\citet{merrill}
